---
title: "example Rmarkdown"
author: "Petr Bouchal"
date: "6/6/2021"
output: html_document
---

This is a toy example of an *Rmarkdown document*. If you are looking at the output, it is a HTML document, which you could put online or email to someone. You can open it in your web browser.

The source is *rendered* into the output HTML by *Knitting* the source document. You could also render a PDF or a Word document - check the options under the Knit button.

The title above is set in the YAML header of the Rmarkdown source document -  the bit between the "---" lines at the very top. You could further tweak the options - check the gear ⚙️ icon in the toolbar.

As you can see, we use *Markdown formatting*.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here are the top lines of a data frame, printed into the output. The lines below between the "fences" are called a *code chunk*.

```{r}
head(mtcars)
```

Below is some calculation. You cannot see the code, but...

```{r, echo=FALSE}
x <- 3*4
```

We multiplied 3 by 4 and the result is `r x`.

The right number shows in the sentence, because we used inline code, between the backticks.

If you have the source Rmarkdown document open in RStudio, you can run code in chunks (Ctrl/Cmd+Enter) or entire chunks by pressing the green play button (▶︎).The output will appear below the code chunk.

Below we do another calculation, where we show the code and have the output appear:

```{r}
5*5
```

(Whether the code and output appear in the output document depends on *chunk options* - click the gear ⚙️ icon on the right of each chunk. There you can set chunk options for each chunk; you can also do it globally in the setup chunk.)

As you can see, emoji and special symbols (čřší) work fine here.

And below is a plot (don't worry about what it shows.)

```{r}
plot(mtcars$mpg, mtcars$wt)
```

You can display your data in tables too - there are many more tools to make tables, but here is a basic one:

```{r}
knitr::kable(head(mtcars))
```

