---
title: "Exploration of Covid-19 vaccinations in CZ"
author: "Petr Bouchal"
date: "6/2/2021"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(gt)
options(scipen = 99)
```

```{r load data}
vd_url <- "https://onemocneni-aktualne.mzcr.cz/api/v2/covid-19/ockovani.csv"
vd_raw <- readr::read_csv(vd_url)
# vd <- readr::read_csv("data-input/vaccines_cz_daily-doses-NUTS3.csv") # or read the data from the downloaded CSV file
```

```{r}
head(vd_raw)
```
```{r}
vd <- vd_raw %>% 
  rename(nuts3 = kraj_nuts_kod,
         nuts3_name = kraj_nazev,
         age_group = vekova_skupina,
         doses_first = prvnich_davek,
         doses_second = druhych_davek,
         doses_total = celkem_davek,
         date = datum,
         vaccine = vakcina) %>% 
  mutate(month = lubridate::floor_date(date, "month"))
```

```{r}
head(vd)
```


```{r, include=FALSE}
vd_reg_month <- vd %>% 
  # there is a row for each day, so let's group by month
  group_by(month, vaccine, nuts3, nuts3_name) %>%
  # and summarise for each of those months
  summarise(doses_total = sum(doses_total)) %>% 
  ungroup() %>% 
  arrange(month) %>% 
  group_by(vaccine, nuts3, nuts3_name) %>% 
  mutate(doses_cum = cumsum(doses_total))
```

## Exploration of raw data

Now let's take a look at cumulative doses by *region* and *vaccine*.

```{r}
vd_reg_month %>% 
  filter(month < "2021-06-01") %>% 
  ggplot() + 
  geom_area(aes(x = month, y = doses_cum, fill = vaccine)) +
  facet_wrap(~ nuts3_name)
```

We are using [daily data (CSV)](`r vd_url`) from the [official Covid data store](https://onemocneni-aktualne.mzcr.cz/api/v2/covid-19).

We could also cut the data by day of the week, not caring about regions and vaccine type:

```{r}
vd_day <- vd %>% 
  group_by(date) %>% 
  summarise(doses_daily = sum(doses_total)) %>% 
  mutate(wday = lubridate::wday(date, label = T, week_start = 1))
```

```{r}
ggplot(vd_day, aes(x = date, y = doses_daily)) +
  geom_col() +
  facet_grid(~wday)
```

## Per capita numbers

But if we want to look at regions and compare them meaningfully, we will need to compare number of doses per capita.

That will require regional population data. One place to get them directly into R is from Eurostat via the [{eurostat}](http://ropengov.github.io/eurostat/) package.

A look through the Eurostat data catalogue as retrieved by `get_eurostat_toc()` suggests the `demo_r_pjanaggr3` table is what we want.

```{r}
pop_all <- eurostat::get_eurostat("demo_r_pjanaggr3")
pop_cz_reg <- pop_all %>% 
  # only keep 2020 data for CZ, for all age groups
  filter(age == "TOTAL", sex == "T",
         # keep lines where NUTS code is NUTS3, like CZ0NN
         str_detect(geo, "CZ0[0-9]{2}"), 
         time == "2020-01-01") %>% 
  rename(pop = values) # name the column unambiguously
```

```{r}
vd_reg_month_pc <- vd_reg_month %>% 
  left_join(pop_cz_reg, by = c("nuts3" = "geo")) %>% 
  mutate(per_cap = doses_cum/pop)
```

Here are the cumulative per capita doses by region and month.

```{r}
ggplot(vd_reg_month_pc, aes(x = month, y = per_cap, fill = vaccine)) +
  geom_col() +
  facet_wrap(~nuts3_name)
```

It would be nice to put these on a map. We can use the [{giscoR}](https://dieghernan.github.io/giscoR/) package to get the boundaries of regions directly into R. (We could quite easily use a shapefile or GeoJSON file via the {sf} package, which is the state-of-the art toolkit for most spatial data in R.)

```{r}
nuts_geo <- giscoR::gisco_get_nuts(country = "CZ", resolution = "03") %>% 
  filter(LEVL_CODE == 3)
```

```{r}
vd_reg_month_pc_latest <- vd_reg_month_pc %>% 
  filter(month == max(month)) %>% # only keep data for latest month, whatever that is
  group_by(nuts3_name, nuts3) %>% # we don't care about breakdown by vaccine type, so summarise
  summarise(per_cap = sum(per_cap))
```

```{r}
nuts_geo %>% 
  left_join(vd_reg_month_pc_latest, by = c("NUTS_ID" = "nuts3")) %>% 
  ggplot(aes(fill = per_cap)) +
  geom_sf(colour = "white") +
  labs(fill = "Vaccine doses per capita") +
  theme_void() +
  scale_fill_viridis_b(labels = scales::number_format(.01), n.breaks = 6) +
  theme(legend.position = "bottom") +
  guides(fill = guide_bins(keywidth = unit(30, "pt")))
```

```{r}
max_per_day <- vd_day %>% filter(doses_daily == max(doses_daily))
max_kraj_per_cap <- vd_reg_month_pc_latest %>% ungroup() %>% filter(per_cap == max(per_cap))
```

The highest nationwide daily count of vaccine doses has been given on `r strftime(max_per_day$date, "%d %B %Y")`.

The most vaccines per person to date have been given in **`r max_kraj_per_cap$kraj_nazev`: `r round(max_kraj_per_cap$per_cap, digits = 3)`** doses per capita.

Here is a table, ordered by per-capita doses given:

```{r, fig.width=6}
gt(vd_reg_month_pc_latest %>% 
     select(nuts3, per_cap) %>% 
     arrange(desc(per_cap)) %>%  
     mutate(per_cap = round(per_cap, 3)) %>% 
     ungroup(),
       auto_align = T) %>% 
  cols_width(nuts3_name ~ 200, per_cap ~ 400) %>% 
  cols_label(nuts3_name = "Kraj", per_cap = "Doses per person") %>% 
  tab_options(table.width = px(100),
                  row.striping.background_color = "lightblue")
```

```{r}
writexl::write_xlsx(vd_reg_month_pc_latest, "data-output/reg_percap_latest.xlsx")
```

# Further exercises

What was the first day on which 10 % of people have had their first jab in each region?

```{r}
vd_reg_day <- vd %>% 
  group_by(date, nuts3, nuts3_name) %>% 
   # summarise multiple columns using the same summary function
  summarise(across(starts_with("doses_"), sum))
```

```{r}
threshold <- 0.1

vd_reg_10pc <- vd_reg_day %>% 
  ungroup() %>% 
  # order by date so we get the cumulative sums riht
  arrange(date) %>% 
  # group by region as we need to calculate cumulative sums across time
  # for each region
  group_by(nuts3, nuts3_name) %>% 
  # do the cumulative sum
  mutate(doses_first_cum = cumsum(doses_first)) %>% 
  # add population table, with nuts3 column matching geo column
  left_join(pop_cz_reg, by = c(nuts3 = "geo")) %>% 
  # calculate per capita
  mutate(doses_first_cum_pc = doses_first_cum/pop) %>% 
  # only keep days when the per capita was above threshold
  filter(doses_first_cum_pc > threshold) %>% 
  # find earliest such date for each region
  group_by(nuts3, nuts3_name) %>% 
  filter(date == min(date)) %>% 
  ungroup() %>% 
  # make regions appear in order of the date when they first hit those 10%
  mutate(nuts3_name = fct_reorder(nuts3_name, date) %>% fct_rev())
  
```

```{r}
ggplot(vd_reg_10pc, aes(x = date, y = nuts3_name)) +
  geom_point() +
  geom_label(aes(label = strftime(date, "%d %b")), 
             colour = "darkgrey", nudge_x = 2) +
  theme_minimal() +
  labs(title = str_glue("Day when each region hit {scales::percent(threshold)} of population with at least first dose"),
       x = NULL, y = NULL)
```
